---
title: "Merge gtex summary statistics"
output: html_document
---

Let's try merging some summary statistics for different tissues.

```{r}
#get a list of all 001 files 
filelist001 = Sys.glob("data/sumstats/*001*")
#set up empty data frame
df=list()
for(i in 1:length(filelist001)){
  df[[i]] = read.table(filelist001[[i]],header=TRUE,stringsAsFactors=FALSE)
}
```

Take a look at the dataframes
```{r}
  head(df[[1]])
  head(df[[2]])
```

Get tissue names and add them as a column
```{r}
get_tissue_name=function(filename){
  suffix = strsplit(filename,c("_sumstats_"))[[1]][2]
  return(strsplit(suffix,".txt.gz")[[1]][1])
}
get_tissue_name(filelist001[1])
tissuenames= unlist(lapply(filelist001,get_tissue_name))
for(i in 1:length(df)){df[[i]]$tissue = tissuenames[i]}
```



Now take a look at dplyr. 
Handy cheatsheet at 
http://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf
```{r}
library(dplyr)
library(tidyr)
library(magrittr)
```

Merge our data into a single big datafram
```{r}
df = rbind_all(df)
dim(df)
```


We can use mutate to add a column of t statistics
```{r}
head(mutate(df,t= betahat.geno/sebetahat.geno))
```

Or we can use the pipe operator %>% to do this (useful for stringing multiple commands together).
 Note that x %>% f(y) is the same as f(x,y).
That is x becomes the first argument to the function it is being piped to.
```{r}
df %>% mutate(t= betahat.geno/sebetahat.geno)
```

Note that this doesn't actually change df. To return the results of the pipe back to df we use
```{r}
df %<>% mutate(t= betahat.geno/sebetahat.geno)
df
```


We can use unite to make the (gene,snp) combination into a single column, eqtl
```{r}
df %>% unite(eqtl, gene, snp,sep="::") 
```


We can group the observations for each eqtl and take the max of the absolute value of the t statistic
```{r}
eqtl_scores = df %>% 
              unite(eqtl, gene, snp,sep="::") %>% 
              group_by(eqtl) %>% 
              summarize(tmax=max(abs(t)))
head(eqtl_scores)
```

We can arrange them by the value of this statistic
```{r}
eqtl_scores %>% arrange(desc(tmax))
```

We can undo the unite to separate back into gene and SNP. Note the use of the %<>% operator from the magrittr library
to assign result back to eqtl_scores.
```{r}
eqtl_scores %<>% separate(eqtl,into=c("gene","snp"),sep="::")
```

We can find the best eqtl in each gene
```{r}
top_eqtl = eqtl_scores %>% group_by(gene) %>% top_n(1,tmax)
```

Now we can filter the original dataframe to only contain the best eQTL in each gene, and add the tmax column
using an inner_join
```{r}
test= df %>% inner_join(top_eqtl)
```

```{r}
tstats = select(test, gene, snp, tissue,t)
```

Finally we can spread out the tissue t stats into "wide" format using spread
```{r}
tstats %<>% spread(tissue,t)
tstats
```



